---
title: "I Watch Therefore I Am: Exploratory Data Analysis of Movie Ratings"
author: "Jack McMorrow, Aditya Kumar, Anthony C. Okoye"
date: "2022-11-03"
output:
  rmdformats::readthedown:
      toc_float: true
      toc_depth: 3
      number_sections: true
      code_folding: hide
      includes:
        before_body: header.tmphtml
---

```{r init, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F, echo = F)
options(scientific=T, digits = 3) 
```

```{r importing data and preprocessing, include=FALSE}
rm(list=ls())
library(tidyverse)


movies <- read_csv("movies.csv")
ratings <- read_csv("ratings.csv")

library(data.table)
movies <- data.table(movies)
dummy_movie_genre_expanded <- movies[, list(title,genres = unlist(strsplit(genres, split = "|", fixed = T))), by = movieId]

# remove spaces
dummy_movie_genre_expanded$genres <- str_trim(dummy_movie_genre_expanded$genres)
str(dummy_movie_genre_expanded)

# joining datasets
movie_ratings_expanded <- full_join(dummy_movie_genre_expanded, ratings, by = "movieId")

# Creating New Variables - To perform 2 sample t-tests for comparing each Genre, with the rest of the Genre.
dummy_movie_wide <- movies %>%
  mutate(
    Adventure = if_else(grepl("Adventure", movies$genres), 1, 0),
    Animation = if_else(grepl("Animation", movies$genres), 1, 0),
    Children = if_else(grepl("Children", movies$genres), 1, 0),
    Comedy = if_else(grepl("Comedy", movies$genres), 1, 0),
    Fantasy = if_else(grepl("Fantasy", movies$genres), 1, 0),
    Romance = if_else(grepl("Romance", movies$genres), 1, 0),
    Drama = if_else(grepl("Drama", movies$genres), 1, 0),
    Action = if_else(grepl("Action", movies$genres), 1, 0),
    Crime = if_else(grepl("Crime", movies$genres), 1, 0),
    Thriller = if_else(grepl("Thriller", movies$genres), 1, 0),
    Horror = if_else(grepl("Horror", movies$genres), 1, 0),
    Mystery = if_else(grepl("Mystery", movies$genres), 1, 0),
    `Sci-Fi` = if_else(grepl("Sci-Fi", movies$genres), 1, 0),
    IMAX = if_else(grepl("IMAX", movies$genres), 1, 0),
    War = if_else(grepl("War", movies$genres), 1, 0),
    Musical = if_else(grepl("Musical", movies$genres), 1, 0),
    Documentary = if_else(grepl("Documentary", movies$genres), 1, 0),
    Western = if_else(grepl("Western", movies$genres), 1, 0),
    `Film-Noir` = if_else(grepl("Film-Noir", movies$genres), 1, 0),
    No_genre_listed = if_else(grepl("(no genres listed)", movies$genres), 1, 0)
  )

# Adding year
movie_ratings_expanded$year <- str_sub(movie_ratings_expanded$title, start = nchar(movie_ratings_expanded$title)-4, end = nchar(movie_ratings_expanded$title)-1) %>% str_trim()
movie_ratings_expanded$year <- as.factor(movie_ratings_expanded$year)

# Calulating movie by genre - Used later for graph
rating_by_genre <- movie_ratings_expanded %>%
  group_by(genres) %>%
  summarize(
    Count = n_distinct(userId),
    Mean = mean(rating, na.rm=T),
    SD = sd(rating, na.rm=T),
    Min = min(rating, na.rm=T),
    Max = max(rating, na.rm=T)
  )
rating_by_genre <- rating_by_genre[order(rating_by_genre$Mean, decreasing = TRUE),]

# Calculating rating by year - Used later for graph
rating_by_year <- movie_ratings_expanded %>%
  group_by(year) %>%
  summarize(
    Count = n_distinct(userId),
    Mean = mean(rating, na.rm=T),
    SD = sd(rating, na.rm=T),
    Min = min(rating, na.rm=T),
    Max = max(rating, na.rm=T)
  )
rating_by_year <- rating_by_year[order(rating_by_year$year),]

# Confidence Interval table - Used later for graph
conf_interval <- data.frame(genre = c("Adventure", "Animation", "Children", "Comedy", "Fantasy", "Romance", "Drama", "Action", "Crime", "Thriller", "Horror", "War", "Musical", "Documentary", "Western", "Film-Noir", "Sci-Fi", "Mystery"),
                            mean = c(3.20, 3.41, 3.09, 3.09, 3.18, 3.23, 3.28, 3.11, 3.24, 3.13, 2.83, 3.39, 3.23, 3.48, 3.25, 3.41, 3.10, 3.21),
                            lower = c(3.16, 3.28, 3.02, 3.07, 3.12, 3.20, 3.26, 3.08, 3.20, 3.10, 2.77, 3.33, 3.15, 3.40, 3.15, 3.31, 3.04, 3.15),
                            upper = c(3.25, 3.48, 3.16, 3.12, 3.24, 3.27, 3.31, 3.15, 3.28, 3.16, 2.89, 3.46, 3.30, 3.56, 3.36, 3.51, 3.15, 3.27))


# Now we can begin writing - Now we can copy-paste our code from the master sheet to input our graphs in our paper.
# Feel free to add more above.

# BEGIN WRITING
```    
   
    
# Background   

For a long time, we’ve liked to engross ourselves in fables and stories for ranging purposes. Cinema has become the modern medium to achieve this. From documentaries to action thrillers, it has become an integral part of any community; movies mirror what we believe and help redefine the narrative on how we see others in the world. Whether they’re utilized as a medium to relax after work or school, address key issues to drive social change, or therapeutically to tackle mental health, movies drive and shape our societies.   

Data science can be a powerful tool for us to better contextualize our movie-watching experience. Can we quantitatively explain what kinds of movies might be better for us to watch than others? Is there data available in the world to address this? We decided to take on this challenge by using movie ratings to try to see what kinds of movies might be considered better than others.    

# Dataset   

(GroupLens)[https://grouplens.org/datasets/movielens/] (Department of Computer Science and Engineering, the University of Minnesota, 2022) is a research lab, which among other things, specializes in a movie recommendation system through MovieLens, that utilizes IMDB user ratings to recommend movies to its users. Using two datasets, one with 10,329 movies, identified by a unique movieID, and their genres, and another with 105,339 IMDB ratings (scale of 5) across 668 Users (selected randomly), identified by a unique UserId, between March 1996 and September 2018, we attempt to perform exciting exploratory analysis of movie ratings, genres, and user preference to gain valuable insights.   

Since our data existed in two separate files, we needed to create a single set to perform analysis.   

## Initial assumptions about the Data   

As is always the case with Data Analysis, we too had few initial assumption about the dataset we were working with. We needed these assumptions to frame some of the SMART questions we wanted to answer. These assumptions can be found below:   

* All movies had associated genres.   
* All movies had only one associated genre.   
* The number of user ratings were evenly spread out across movies, and hence genres.   
* Movies present were released only during the specified time range.   

