---
title: "Movie Ratings Group Project"
author: "Group 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing Data
```{r}
rm(list=ls())
library(tidyverse)

movies <- read_csv("movies.csv")
ratings <- read_csv("ratings.csv")

str(movies)
str(ratings)
```

## Data Proprocessing

Expanding Movies Dataset
```{r}
library(data.table)
movies <- data.table(movies)
movies_expanded <- movies[ , list(genres = unlist(strsplit(genres, split = "|", fixed = T))), by = movieId]

# remove spaces
movies_expanded$genres <- str_trim(movies_expanded$genres)
str(movies_expanded)
```

Checking Movies with only one genre listed
```{r}
movie_count_one <- movies_expanded %>%
  group_by(movieId) %>%
  summarize(
    count = n()
  ) %>%
  filter(count == 1)
movie_count_one
```

3,041 movies have only one genre listed (29.4%)
*Might be able to perform a chi-squared goodness of fit test to see if there is any bias between movies with only one genre listed versus those with multiple genres*


Joining Datasets
```{r}
movies_drop <- movies %>% select(-3)
movies_join <- left_join(movies_expanded, movies_drop, by = "movieId")
data <- full_join(movies_join, ratings, by = "movieId")
```

Subsetting Movies with one genre
```{r}
one_genre <- inner_join(movies_join, movie_count_one)
one_genre
```

Finding proportion of genres
```{r}
one_genre_prop <- one_genre %>%
  group_by(genres) %>%
  summarize(
    count = n(),
    prop = n() / 3041
  )

one_genre_prop <- one_genre_prop[order(one_genre_prop$prop, decreasing = T),]
one_genre_prop
```

```{r}
all_movies_prop <- movies_join %>%
  group_by(genres) %>%
  summarize(
    count = n(),
    prop = n() / 10329
  )

all_movies_prop <- all_movies_prop[order(all_movies_prop$prop, decreasing = T),]
all_movies_prop
```

One genre category is missing the IMAX category
```{r}
prop_join <- full_join(one_genre_prop, all_movies_prop, by = "genres")
prop_join <- rename(prop_join,
       count_one = count.x,
       prop_one = prop.x,
       count_all = count.y,
       prop_all = prop.y)
prop_join[is.na(prop_join)] = 0
prop_join
```

Doing a Chi-squared Goodness of fit test to see if proportions of one genre are different from all of the genres
```{r}
# chisq.test(prop_join$count_one, p = prop_join$prop_all)
```

Chi-squared test does not work because the proportion of genres does not equal 1. We'll have to think of another way to see if these two stats are significantly different

```{r}
unique(movies_join$genres)
sum(is.na(movies_join$genres))
```

20 different genres

Some issues with the final dataset:
* There are multiple genres per movie and multiple reviews per movie. This means the review ratings are duplicated in the final dataset in movies with multiple genres after they have been expanded. We might want to join the datasets with all genres listed in one column instead.
* Overlap of genres between different movies might cause some issues.
* The dataset is very long. We might want to subset it to those movies with just one genre (which would address the issue above), or perhaps narrow it down to a certain time frame.
* The timestamp variable is being interpreted as numeric. We will need to figure out how to fix this.

*Extract year from movie titles*

Columns to make:
* Average user ratings per movie
* Number of ratings per movie
* Year movie is made

```{r}
genre_list <- list(unique(movies_join$genres))
for (g in genre_list) {
  movie_wide <- movies %>%
    mutate(
      g = if_else(grepl(g, movies$genres), 1, 0)
    )
}

str(movie_wide)
unique(movies_join$genres)
```

Creating New Variables
```{r}
movie_wide <- movies %>%
  mutate(
    Adventure = if_else(grepl("Adventure", movies$genres), 1, 0),
    Animation = if_else(grepl("Animation", movies$genres), 1, 0),
    Children = if_else(grepl("Children", movies$genres), 1, 0),
    Comedy = if_else(grepl("Comedy", movies$genres), 1, 0),
    Fantasy = if_else(grepl("Fantasy", movies$genres), 1, 0),
    Romance = if_else(grepl("Romance", movies$genres), 1, 0),
    Drama = if_else(grepl("Drama", movies$genres), 1, 0),
    Action = if_else(grepl("Action", movies$genres), 1, 0),
    Crime = if_else(grepl("Crime", movies$genres), 1, 0),
    Thriller = if_else(grepl("Thriller", movies$genres), 1, 0),
    Horror = if_else(grepl("Horror", movies$genres), 1, 0),
    Mystery = if_else(grepl("Mystery", movies$genres), 1, 0),
    `Sci-Fi` = if_else(grepl("Sci-Fi", movies$genres), 1, 0),
    IMAX = if_else(grepl("IMAX", movies$genres), 1, 0),
    War = if_else(grepl("Adventure", movies$genres), 1, 0),
    Musical = if_else(grepl("Musical", movies$genres), 1, 0),
    Documentary = if_else(grepl("Documentary", movies$genres), 1, 0),
    Western = if_else(grepl("Western", movies$genres), 1, 0),
    `Film-Noir` = if_else(grepl("Film-Noir", movies$genres), 1, 0),
    No_genre_listed = if_else(grepl("(no genres listed)", movies$genres), 1, 0)
  )
str(movie_wide)
```

```{r}
sapply(movie_wide, mean)
```


```{r}
rating_mean <- ratings %>%
  group_by(movieId) %>%
  summarize(
    mean_rating = mean(rating)
  )

movies_rating <- left_join(movie_wide, rating_mean, by = 'movieId')
movies_rating
```

Expanding with Movie Ratings
```{r}
library(data.table)
movies_rating <- data.table(movies_rating)
movies_rating_expanded <- movies_rating[ , list(genres = unlist(strsplit(genres, split = "|", fixed = T))), by = movieId]

# remove spaces
movies_rating_expanded$genres <- str_trim(movies_rating_expanded$genres)

# rejoining tables
movies_rating <- select(movies_rating, -genres)
movies_rating_expanded <- full_join(movies_rating_expanded, movies_rating, by = 'movieId')
str(movies_rating_expanded)
```

Creating Year Variable
```{r}
movies_rating_expanded$year <- str_sub(movies_rating_expanded$title, start = nchar(movies_rating_expanded$title)-4, end = nchar(movies_rating_expanded$title)-1) %>% str_trim()
movies_rating_expanded$year <- as.factor(movies_rating_expanded$year)

movie_wide$year <- str_sub(movie_wide$title, start = nchar(movie_wide$title)-4, end = nchar(movie_wide$title)-1) %>% str_trim()
movie_wide$year <- as.factor(movie_wide$year)

str(movie_wide)
```

## Data Visualization

```{r}
rating_by_genre <- movies_rating_expanded %>%
  group_by(genres) %>%
  summarize(
    Count = n(),
    Mean = mean(mean_rating, na.rm=T),
    SD = sd(mean_rating, na.rm=T),
    Min = min(mean_rating, na.rm=T),
    Max = max(mean_rating, na.rm=T)
  )

rating_by_genre <- rating_by_genre[order(rating_by_genre$Mean, decreasing = TRUE),]
rating_by_genre
```

```{r}
rating_by_genre %>%
  ggplot(
    aes(
      x = fct_reorder(genres, Count, .desc = T),
      y = Mean
    )
  ) +
  geom_col(fill = "lightblue", color = "black") +
  labs(
    title = "User Movie Rating by Genre",
    x = "Genre",
    y = "Average User Rating (0-5)"
  ) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Calculating Rating by Year
```{r}
rating_by_year <- movies_rating_expanded %>%
  group_by(year) %>%
  summarize(
    Count = n(),
    Mean = mean(mean_rating, na.rm=T),
    SD = sd(mean_rating, na.rm=T),
    Min = min(mean_rating, na.rm=T),
    Max = max(mean_rating, na.rm=T)
  )

rating_by_year <- rating_by_year[order(rating_by_year$year),]
rating_by_year
```

```{r}
rating_by_year %>%
  ggplot(
    aes(
      x = year,
      y = Mean
    )
  ) + 
  geom_col(fill = "lightblue") +
   labs(
    title = "User Movie Rating by Year",
    x = "Year",
    y = "Average User Rating (0-5)"
  ) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
```

## Statistical Testing
Testing Adventure
```{r}
Adventure <- movies_rating %>% filter(Adventure == 1)
Not_adventure <- movies_rating %>% filter(Adventure == 0)

t.test(Adventure$mean_rating)
t.test(Adventure$mean_rating, Not_adventure$mean_rating)
```

95% Confidence Interval for Adventure Movies: 3.16 - 3.25.

Two-sample t-test has a p-value of 0.2, so we cannot reject null hypothesis that Adventure movies score higher than all other movies.


